Forecasting biodiversity in breeding birds using best practices

### 2017-12-08 manuscript/draft.Rmd
Update draft.Rmd

Forecasting is an emerging imperative in ecology; as such, the field needs to
develop and follow best practices for conducting and evaluating ecological
forecasts [@clark2001]. We have used a number of these practices (Box 1) in a
single study that builds and evaluates forecasts of biodiversity in the form of
species richness. The results of this effort are both promising and humbling.
When comparing forecasts across sites, many different approaches to forecasting
produce reasonable forecasts (Figure 3). If a site is predicted to have a high
number of species in the future, relative to other sites, it generally does.
However, none of the methods evaluated reliably determined how site-level
richness changes over time (Figure 6), which is generally the stated purpose of
these forecasts. As a result, baseline models, which did not attempt to
anticipate changes in richness over time, generally provided the best forecasts
for future biodiversity. While this study is restricted to breeding birds in
simple baselines. Since site-level richness is relatively stable in so many
systems [@dornelas2014], spatial validation will not be enough: a model can have
high accuracy across spatial gradients without being able to predict changes
over time. This gap between spatial and temporal accuracy is known to be
important for species-level predictions [@rapacciuolo2012; @oedekoven2017]; our
results indicate that it is substantial for higher-level patterns like richness
as well. SDMs' poor temporal predictions are particularly sobering, as these
models have been one of the main foundations for estimates of the predicted loss
of biodiversity to climate change over the past decade or so [@thomas2004;
@thuiller2011; @urban2015]. One could argue that these conclusions are unfair to
SDMs and related methods, which are generally used to predict large directional
changes over multi-decade timescales rather than fluctuations within a single
decade. It is certainly possible that these models' relative performance would
improve over time or with larger shifts in richness, and future research should
explore this possibility.

@@

When comparing different methods for forecasting our results demonstrate the
importance of considering uncertainty [Box 1; @clark2001; @dietze2016]. Previous
comparisons between stacked SDMs and macroecological models reported that the
methods yielded equivalent results for forecasting diversity [@algar2009;
@distler2015]. While our results support this equivalence for point estimates,
they also show that stacked SDMs dramatically underestimate the range of
possible outcomes; after ten years, more than a third of the observed richness
values fell outside the stacked SDMs' 95% prediction intervals. Consistent with
@harris2015 and @warton2015, we found that JSDMs' wider prediction intervals
enabled them to avoid this problem. Macroecological models appear to share this
advantage, while being considerably easier to implement.

>>>>

Forecasting is an emerging imperative in ecology; as such, the field needs to
develop and follow best practices for conducting and evaluating ecological
forecasts [@clark2001]. We have used a number of these practices (Box 1) in a
single study that builds and evaluates forecasts of biodiversity in the form of
species richness. The results of this effort are both promising and humbling.
When comparing forecasts across sites, many different approaches to forecasting
produce reasonable forecasts (Figure 3). If a site is predicted to have a high
number of species in the future, relative to other sites, it generally does.
However, none of the methods evaluated reliably determined how site-level
richness changes over time (Figure 6), which is generally the stated purpose of
these forecasts. As a result, baseline models, which did not attempt to
anticipate changes in richness over time, generally provided the best forecasts
for future biodiversity. While this study is restricted to breeding birds in
simple baselines. Since site-level richness is relatively stable in so many
systems [@dornelas2014], spatial validation will not be enough: a model can have
high accuracy across spatial gradients without being able to predict changes
over time. This gap between spatial and temporal accuracy is known to be
important for species-level predictions [@rapacciuolo2012; @oedekoven2017]; our
results indicate that it is substantial for higher-level patterns like richness
as well.

SDMs' poor temporal predictions are particularly sobering, as these models have
been one of the main foundations for estimates of the predicted loss of
biodiversity to climate change over the past decade or so [@thomas2004;
@thuiller2011; @urban2015]. One could argue that these conclusions are unfair to
SDMs and related methods, which are generally used to predict large directional
changes over multi-decade timescales rather than fluctuations within a single
decade. It is certainly possible that these models' relative performance would
improve over time or with larger shifts in richness, and future research should
explore this possibility. Nevertheless, our results have important implications
for ecologists' long-term predictions regarding large changes in species
richness, for two reasons. First, after controlling for observer effects, we
found no evidence that these models have any ability to track changes in
richness over time at all (Figure 6). Given that prediction error steadily
increases over time (Figure 5), we expect that long-term SDM-based forecasts
will generally be at least as poor as decade-scale ones.

@@

When comparing different methods for forecasting our results demonstrate the
importance of considering uncertainty [Box 1; @clark2001; @dietze2016]. Previous
comparisons between stacked SDMs and macroecological models reported that the
methods yielded equivalent results for forecasting diversity [@algar2009;
@distler2015]. While our results support this equivalence for point estimates,
they also show that stacked SDMs dramatically underestimate the range of
possible outcomes; after ten years, more than a third of the observed richness
values fell outside the stacked SDMs' 95% prediction intervals. Consistent with
@harris2015 and @warton2015, we found that JSDMs' wider prediction intervals
mostly enabled them to avoid this problem. Macroecological models appear to share 
this advantage, while being considerably easier to implement.
